---
title: "Time_Series"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(corrplot)
library(plyr)
library(reshape2)
```

##This RScripts plots various phenotypes against time of measurement

load the csv of phenotype data
```{r load, echo=FALSE}
setwd("~/Projects/DO")
phenotype=read.csv("data/DO_phenotype_data_clean.csv")
desc=read.csv("data/svenson850_phenotype_descr.csv")
indBW=grep("BW.*",names(phenotype))
BW.dat=phenotype[,c(4:8, indBW)]
BW.dat$sexDiet=paste(BW.dat$Sex, BW.dat$Diet, sep="_")
#look
summary(BW.dat)[,1:10]
```

Reshape data for ggplotting

```{r format}
BWmelt=melt(BW.dat, id.vars = c("Sample","Sex","Gen","parity", "Diet", "sexDiet"))
BWmelt.clean=na.omit(BWmelt)
head(BWmelt.clean)
```

##Plot out 
Progression of Body Weight factored by sex and diet
```{r }
ggplot(BWmelt.clean, aes(x=as.numeric(variable)+3, y=as.numeric(value), col=sexDiet, group=sexDiet))+
  geom_smooth()+
  geom_jitter(width=0.5, size=0.01)+
  ggtitle("Body weight | Sex and Diet")+
  theme_classic()+
  labs(y="grams", x="Age (wks)")
```

Progression of Body Weight factored by diet
```{r, echo=FALSE}
ggplot(BWmelt, aes(x=as.numeric(variable)+3, y=as.numeric(value), col=Diet, group=(Diet)))+
  geom_smooth()+
  geom_jitter(width=0.5, size=0.01)+
  ggtitle("Body weight | Diet")+
  theme_classic()+
  labs(y="grams", x="Age (wks)")
```
Progression of Body Weight factored by sex
```{r, echo=FALSE}
ggplot(BWmelt, aes(x=as.numeric(variable)+3, y=as.numeric(value), col=Sex, group=Sex))+
  geom_smooth()+
  geom_jitter(width=0.5, size=0.01)+
  ggtitle("Body weight | Sex")+
  theme_classic()+
  labs(y="grams", x="Age (wks)")

```

Progression of Body Weight factored by generation
```{r, echo=FALSE}
ggplot(BWmelt, aes(x=as.numeric(variable)+3, y=as.numeric(value), col=Gen, group=Gen))+
  geom_smooth()+
  geom_jitter(width=0.5, size=0.01)+
  ggtitle("Body weight | Gen")+
  theme_classic()+
  labs(y="grams", x="Age (wks)")
```

##Other phenotypes that are measured at 2 time points
```{r}
#find all the phenotypes that have 1 and 2 at end, those are temporal phenotypes of interest
indtemp=c(grep("*1",names(phenotype), value=F),grep("*2",names(phenotype), value=F))
subset=phenotype[,c(5,8,sort(indtemp)[-1:-2])]
rownames(subset)=phenotype$Sample
#take out the body weight data, will deal with those later
minus=grep("BW.*", names(subset), value=F)
timDat=subset[,-c(minus)]
summary(timDat)[,1:10]
```

Melt and reformat the data
```{r, echo=FALSE}
#reformat data and delet the 1 and 2
meltDat=melt(timDat, id.vars=c("Sex","Diet"))
#delete the 1 and 2 at end to make phenotypes same
#remove all [.] first
meltDat$variable=gsub("[.]","", meltDat$variable)
#gsub back in ".1" and ".2"
meltDat$variable=gsub("2",".2", meltDat$variable)
meltDat$variable=gsub("1",".1", meltDat$variable)
#split by [.], then unsplit and add back into frame
ls=unlist(strsplit(meltDat$variable,"[.]"))
ls.mat=matrix(ls, nrow=2)
meltDat$variable=ls.mat[1,]
meltDat$week=ls.mat[2,]

#split data into a list of 49 phenotypes
desc$name=gsub("[.]", "", desc$name)
meltList=split(meltDat, meltDat$variable)
attributes(meltList)
```

##Plot out
```{r, include=FALSE, warning=FALSE}
a=desc$weeks[grep(pattern=paste(meltList$percFat$variable, "*", sep=""), desc$name)][1]
b=desc$weeks[grep(pattern=paste(meltList$percFat$variable, "*", sep=""), desc[,1])][2]
p1=ggplot(na.omit(meltList$percFat), aes(y=as.numeric(value), x=week, group=Diet, col=Diet, warnings=F))+
  geom_smooth()+
  geom_jitter(width=0.08, size=0.1)+
  theme_classic()+
  labs(y="value")+
  scale_x_discrete(labels=c(a,b))+
  theme(legend.title=element_blank(), 
        plot.title=element_text(size=9),
        legend.position="top") +
  ggtitle("percFat | Diet")

p2=ggplot(na.omit(meltList$percFat), aes(y=as.numeric(value), x=week, group=Sex, col=Sex, warnings=F))+
  geom_smooth()+
  geom_jitter(width=0.08, size=0.1)+
  theme_classic()+
  labs(y="value")+
  scale_x_discrete(labels=c(a,b))+
  theme(legend.title=element_blank(), 
        plot.title=element_text(size=9),
        legend.position="top") +
  ggtitle("percFat | Sex")


```


Plot phenotype factored by diet and then sex (Scroll pass the warnings)
```{r, warning=FALSE}
########################################################
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
multiplot(p1,p2, cols=2)
```

