---
title: "Mediation_Scans"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Looking for Candidate Genes by mediating RNA expressions on the Svenson DO phenotypes

Load 
```{r load, warning=FALSE}
setwd("~/Projects/DO")
if(!require(c("devtools", "RcppEigen"))) {
  install.packages(c("devtools", "RcppEigen"),
  repos = 'http://cran.rstudio.com/')
}
devtools::install_github("simecek/intermediate")
library(intermediate)
load("data/DO478_ExprData_4eQTLMapping.Rdata")
load("data/Svenson_4eQTLMapping.Rdata", verbose=T)
load("results/7-26-16/Svenson_pheno_intdiet.RData", verbose=T)
load("results/7-26-16/Svenson_pheno_intsex.RData", verbose=T)
load("results/7-26-16/Svenson_pheno_add.RData", verbose=T)
```

## Preparing the arguments
```{r prep}
##update the mediation scan function to accomodate interactive covars
mediation.scan <-
  function(target,
           mediator,
           annotation,
           qtl.geno,
           addcovar = NULL,
           intcovar = NULL,
           method = c("double-lod-diff", "ignore", "lod-diff", "lod-ratio"),
           verbose = TRUE)
  {
    LL <- function(y, X) {
      -length(y) / 2 * log10(sum(qr.resid(qr(cbind(
        X, 1
      )), y) ^ 2))
    }
    stopifnot(NROW(target) == NROW(mediator))
    stopifnot(NROW(annotation) == NCOL(mediator))
    stopifnot(NROW(qtl.geno) == NROW(target))
    stopifnot(is.null(addcovar) | NROW(target) == NROW(addcovar))
    stopifnot(!any(is.na(addcovar)))
    stopifnot(is.null(intcovar) | NROW(target) == NROW(intcovar))
    stopifnot(!any(is.na(intcovar)))
    stopifnot(!any(is.na(qtl.geno)))
    stopifnot(all(is.numeric(target)))
    stopifnot(all(is.numeric(mediator)))
    stopifnot(all(is.numeric(qtl.geno)))
    stopifnot(all(is.numeric(addcovar)))
    stopifnot(all(is.numeric(intcovar)))
    stopifnot(c("CHR", "POS") %in% toupper(names(annotation)))
    method = match.arg(method)
    mediator <- cbind(mediator)
    N <- ncol(mediator)
    if (is.null(addcovar))
      addcovar <- cbind(rep(1, N))
    LOD <- rep(NA, N)
    if (!is.null(intcovar))
      qtl.geno <- cbind(qtl.geno, qtl.geno * as.numeric(intcovar))
    if (method == "double-lod-diff") {
      no.na <- !is.na(target)
      LOD0 <- LL(target[no.na], cbind(addcovar, qtl.geno)[no.na,]) - LL(target[no.na], addcovar[no.na,])
    }
    for (i in 1:N) {
      if (verbose & i %% 1000 == 0)
        print(i)
      no.na <- !is.na(target) & !is.na(mediator[, i])
      loglik0 <-
        LL(target[no.na], cbind(addcovar[no.na,], mediator[no.na,
                                                           i]))
      loglik1 <-
        LL(target[no.na], cbind(addcovar[no.na,], mediator[no.na,
                                                           i], qtl.geno[no.na,]))
      if (method == "ignore" | (method == "double-lod-diff" &
                                all(no.na))) {
        LOD[i] <- loglik1 - loglik0
      }
      else {
        loglik2 <- LL(target[no.na], addcovar[no.na,])
        loglik3 <- LL(target[no.na], cbind(addcovar[no.na,],
                                           qtl.geno[no.na,]))
        if (method == "lod-diff") {
          LOD[i] <- loglik3 - loglik2 - (loglik1 - loglik0)
        }
        else if (method == "double-lod-diff") {
          LOD[i] <- LOD0 - (loglik3 - loglik2 - (loglik1 -
                                                   loglik0))
        }
        else if (method == "lod-ratio") {
          LOD[i] <- (10 ^ loglik1 - 10 ^ loglik0) / (10 ^ loglik3 -
                                                       10 ^ loglik2)
        }
      }
    }
    output <- annotation
    output$LOD <- LOD
    class(output) <- c("mediation", "data.frame")
    return(output)
  }
##

##
#make interactive covariates
#sex
sexcovar=as.matrix(covar[,1])
colnames(sexcovar)=colnames(covar)[1]
#diet
dietcovar=as.matrix(covar[,4])
colnames(dietcovar)=colnames(covar)[4]
#
##Change the 4th name on annot
names(annot)[3:4]=c("chr","pos")

```

##Doing the mediation
```{r med}
#Mediation scan for addtive covariates
med.add = mediation.scan(
  target = phenotype[, "CHOL2"],
  mediator = expr,
  annotation = annot,
  covar = covar,
  #pick out the marker that has the highest LOD peak
  qtl.geno = probs.478[, , maxlod.add$marker[which(maxlod.add[, 4] ==
                                                     "CHOL2")]]
)
plot(med.add)
kplot(med.intadd)



#Mediation scan for interactive diet covariates
med.intdiet = mediation.scan(
  target = phenotype[, "CHOL2"],
  mediator = expr,
  annotation = annot,
  addcovar = covar,
  intcovar = dietcovar,
  #pick out the marker that has the highest LOD peak
  qtl.geno = probs.478[, , maxlod.add$marker[which(maxlod.intdiet[, 4] ==
                                                     "CHOL2")]]
)
plot(med.intdiet)
kplot(med.intdiet)

